
STA130 Capstone Project: Understanding Red Giants

```{r}
library(tidyverse)
library(rhdf5)
```

```{r}
file = "STA130_APOGEE.h5"
```

Research Question 1:
Are red giants with higher intensity measurements harder to measure?

Extracting relevant columns from the dataset:
```{r}
spectra_1 <- file %>% h5read("spectra", index=list(NULL, 1:30000)) %>% t() %>% as_tibble()
spectra_2 <- file %>% h5read("spectra", index=list(NULL, 30001:60000)) %>% t() %>% as_tibble()
spectra_3 <- file %>% h5read("spectra", index=list(NULL, 60001:99705)) %>% t() %>% as_tibble()

snr <- file %>% h5read("snr") %>% as_tibble()
```

Cleaning the data:
```{r}
spectra_means_1 <- apply(spectra_1, 1, mean) %>% as_tibble()
spectra_means_2 <- apply(spectra_2, 1, mean) %>% as_tibble()
spectra_means_3 <- apply(spectra_3, 1, mean) %>% as_tibble()
spectra_means <- rbind(spectra_means_1, spectra_means_2, spectra_means_3)
```

Analysis:
```{r}
lm(snr$value ~ spectra_means$value) %>% summary()
mod = lm(snr$value ~ spectra_means$value)
res = summary(mod)

pt(coef(res)[, 3], mod$df, lower = TRUE)

```
Visualizations:
```{r}
ggplot() + aes(x=spectra_means$value, y=snr$value) + geom_point() + geom_smooth(method="lm", se=FALSE) +
  labs(title="Scatter Plot of Average Intensity Measurements against Snr", x="Average Intensity Measurements", y="Signal-to-noise Ratio")
ggplot() + aes(x=spectra_means$value, y=snr$value) + geom_quantile() + labs(title="Quantile Plot of Average Intensity Measurements against Snr", x="Average Intensity Measurements", y="Signal-to-noise Ratio")
```

Research Question 2:
Does surface gravity vary by a small factor among different red giants?

Extracting the relevant columns:
```{r}
logg <- file %>% h5read("logg") %>% as_tibble()
```

Analysis:
```{r}
logg_original_sample <- logg %>% sample_n(size=10000)

sample_cov <- rep(NA,1000)
for(i in 1:1000){
  logg10000 <- logg_original_sample %>% sample_n(size=10000,replace = TRUE)
  sample_cov[i]<- as.numeric(logg10000 %>% summarise(sd(value))) / as.numeric(logg10000 %>% summarise(mean(value)))
}
logg_cov <- tibble(coeff_of_var = sample_cov)
 
quantile(logg_cov$coeff_of_var,c(0.025,0.975))
```

Visualizations:
```{r}
logg_cov %>% ggplot() + aes(x = coeff_of_var) + 
  geom_histogram(bins=30, colour = 'black', fill = "grey") + 
  geom_vline(aes(xintercept = 0.2236308), color='red') + 
  geom_vline(aes(xintercept = 0.2308553), color='red') +
  labs(title='Histogram of Distribution of Coefficient of Variation over 1000 bootstrap samples
  of size 10000. The red lines show the 95% confidence interval.',
       x='Coefficient of Variation', y='Count')
 
logg_cov %>% ggplot() + aes(x = coeff_of_var) + 
  geom_boxplot() + 
  labs(title='Boxplot of Distribution of Coefficient of Variation over 1000 bootstrap samples 
       of size 10000', 
       x= 'Coefficient of Variation')
```

Question 3:
Is there a correlation between the abundance of carbon on a red giant star and the effective temperature of that star?

Extracting the relevant columns:
```{r}
carbon_abundance <- file %>% h5read("c_h") %>% as_tibble() 
temp_eff <- file %>% h5read("teff") %>% as_tibble()
```

Visualizations:
```{r}
ggplot() + aes(x=carbon_abundance$value, y=temp_eff$value) + geom_point() + 
  geom_smooth(method="lm", se=FALSE) +
  labs(title="Scatter Plot of Effective Temperature against Abundance of Carbon", 
       x='Abundance of Carbon', y='Effective Temperature')
```



